Parameterization of the responses of subarctic European vegetation to key environmental variables for ozone risk assessment

### 2022-04-21 VegOzone.tex
Rephrase decideous tree paragraph

We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations. Canopy height and leaf size are adjusted to local values in the model simulations with the bespoke growing season (GS besp.). The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of over- or underestimation of ozone risk arising from the use of generic parameterizations established from less climatically extreme conditions.
Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. We identify that an increase in cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction. The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018. Due to the shape of $f_\mathrm{light}$, a symmetric variation leads to an asymmetric response in $\mathrm{POD_1}$. We find that an opening of stomata at lower light intensities can cause higher sensitivity to drought conditions. Similarly, a temperature acclimation to cold or subarctic conditions can lead to an amplification of effects by droughts.

The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke temperature parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here. Our bespoke temperature parameterizations suggest considerable underestimations of ozone risk for subarctic species when relying on the generic MM parameterizations.

For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Deciduous trees display the largest variation of ozone uptake among all PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light parameterization is pronounced. With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019, throughout our analyses. However, taking the bespoke GS into account, this is reversed for the subarctic temperature acclimation. This can be explained by more favorable growing conditions in 2019. Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke GS. The temperature acclimation-dependent difference in ozone uptake determined for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum deviations determined from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

@@

From the difference between 2018 and 2019, we infer a substantial interannual variability. The maximum deviation from the generic default parameterizations due to our bespoke temperature and light response functions is of the same order of magnitude as the interannual variability.

>>>>

We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our localized temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.

The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of over- or underestimation of ozone risk arising from the use of MM parameterizations established from less climatically extreme conditions. Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. The different background colors indicate the temperature response function localization. The light response function localization is denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8). Circles indicate simulations with localized GS, canopy height, and leaf dimensions and squares simulations with MM GS. With open symbols the effect on $\mathrm{POD_y}$ of taking SWP into account is displayed.

We identify that an increase in cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction. The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018 (see Section~\ref{subsec:anomalies} for details). Due to the shape of $f_\mathrm{light}$, a symmetric variation leads to an asymmetric response in $\mathrm{POD_1}$. We find that an opening of stomata at lower light intensities will cause a reduction of $\mathrm{POD_1}$ due to drought conditions. This is seen for both $f_\mathrm{T}$ localization in 2018 but not 2019 (compare PPFD0.8 and PPFD1.2 in Fig.~\ref{fig:pody_rel}a, subarctic, 2018, with SWP effect on).

For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Among all PFTs, deciduous trees display the largest sensitivity in $\mathrm{POD_1}$. Compared to MM $\Delta\mathrm{POD_1}=(10-31)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. In particular, the sensitivity to the $f_\mathrm{light}$ is pronounced (extent of $\angle(+,0,-)$). With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019. However, taking the localized GS into account, this is reversed for the subarctic temperature response function. This can be explained by more favorable growing conditions in 2019. Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM GS, but did not or only slightly when using the localized GS. This indicates (frost) drought conditions early in the regular GS in 2018 when the ground was still covered in snow. The difference in $\mathrm{POD_1}$ due to the localization of $f_\mathrm{T}$ for the localized GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum difference due to the $f_\mathrm{light}$ localization for the PPFD0.8 (superscript) and PPFD1.2 (subscript) amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

@@

From the difference between 2018 and 2019, we conclude that the interannual variability of ozone dose is non-negligible. The maximum difference in $\mathrm{POD_1}$ between simulations with the MM parameterizations and the localized temperature and light response functions is of the same order of magnitude as the interannual difference.
The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years, but the predicted ozone uptake for the localized $f_\mathrm{T}$ is always larger than for the MM parameterizations. This suggests underestimation of ozone risk for subarctic species when using the MM parameterizations.
